\documentclass{report}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[hidelinks]{hyperref}

\input{macros}

\title{
  \booktitle \\
  \ \\
  Errata List
}

\begin{document}

\maketitle

\begin{enumerate}

\item Page 62, Exercise 3.5.4.
  The hint should read ``Let $h(x) = A - x$; \ldots'' instead of ``Let $h(x) = B - x$; \ldots''.
  (Confirmed by Dr. Jech)

\item Page 101, Exercise 5.2.3.
  This means that that the countable dense subset is dense in $P$ rather than simply dense with respect to itself.
  This has to be the case because, if were merely dense with respect to itself, then \emph{any} larger linearly ordered set containing the subset would have the same property so that it is impossible to put a bound on the size of that set (Confirmed)

\item Page 114, Exercise 6.3.5 part (c).
  This should read ``\ldots, then $f[A] \in V_\w$.'' instead of ``\ldots, then $f[X] \in V_\w$.'' since $X$ has not been previously defined.
  (Unconfirmed)

\item Page 140, bottom.
  The reference to Assumption~1.7 in Chapter~4 should really be Assumption~1.8.
  (Unconfirmed)

\item Page 142, bottom.
  The reference to Theorem~4.4 in Chapter~7 should be Theorem~4.4 in Chapter~6.
  (Unconfirmed)

\item Page 143 top.
  In the proof of Theorem~8.1.13 when showing that (c) implies (a) the sentence, ``Let $F$ be the system of all functions $f$ for which $\dom{f} \ss S$ and $f(X) \in X$ holds for any $X \in S$.'' should be, ``\ldots and $f(X) \in X$ holds for any $X \in \dom{f}$.''
  This is because, if $\dom{f} \pss S$, then there is an $X \in S$ where $X \notin \dom{f}$ so that $f(X)$ is not defined.
  (Unconfirmed)

\item Page 143, middle.
  In the proof of Theorem~8.1.14 there are two times in the first paragraph where statements are made for all or for some $\x \in A$.
  These should be $\x < \l$.
  (Unconfirmed)

\item Page 158, bottom.
  Near the end of the first paragraph of the proof of Theorem~9.1.7 (K{\"o}nig's Theorem), the sentence, ``If $i_x \neq i_y = i$, then $a_i = d_i \notin A$ while $b_i = y \in A$.'' should be, ``If $i_x \neq i_y = i$, then $a_i = d_i \notin A_i$ while $b_i = y \in A_i$.''
  (Unconfirmed)

\item Page 160, Exercise 9.1.11.
  The special case mentioned in the hint should be $\parens{\k \cdot \l}^\mu = \k^\mu \cdot \l^\mu$.
  This is evidenced by the fact that $\parens{\k^\mu}^\l = \parens{\k^\l}^\mu$ does not make sense in the context of the problem (i.e. it is not a special case) as well as by the fact that it is not part of Theorem~5.1.7.
  (Unconfirmed)

\end{enumerate}

\end{document}
